---
title: "DE Analysis Report (Protein): `r params$title`"
subtitle: "`r params$subtitle`"
author: "`r params$author`"
format: 
  html:
    fig-width: 12
    fig-height: 10
    toc: true
    toc-title: 'DE Analysis'
    title-block-banner: '#1b2944'
    title-block-banner-color: '#f79646'

execute: 
  echo: false
  warning: false
embed-resources: true
theme: 
 - layout/brand 
 - layout/custom.scss

params:
  description: ""
  title: ""
  subtitle: ""
  author: "" 
  formula: ""
  comparisons: ""
  contrast: ""
  input_file: ""
  design_file: ""
  FC_thr: 1
  adjpval_thr: 0.05
  Proteotypic: TRUE
  pep_per_prot: 3
  nNonZero: 30
  aggr_method: "medianPolish"
  normalization: "quantiles"
  filtering_contaminant: ""
  contaminat_str: ""
  folder_prj: ""
  ensembl_annotation: ""
  ensembl_col: ""
  confounder_list: ""
  PCA_comparison: "Group"
  quantitative_features: "Precursor.Quantity"
  filtPerGroup: ''
  mbr: FALSE
  wildstr_run: "CMB-"
  DIANN_ver2: FALSE 
  keep_design_order: FALSE
  comparison_label: ""
---

```{r input}
 #| output: false
source('utils_function.R')

required_packages = c("QFeatures","msqrob2","readxl","SummarizedExperiment","visdat","ggrepel", "dplyr","tidyverse","magrittr","here","tidyr","plotly","DT","tibble","MSnbase",
                    "factoextra","hrbrthemes","gridExtra","UpSetR",'logger', 'scales','pander','heatmaply','GGally','arrow','yaml')
#print(required_packages)
check_dependencies(required_packages)

```

## Experiment Description 

`r params$description`

```{r check_parameter}

log_threshold(INFO)
log_appender(appender_console)
log_appender(appender_file("logfile_protein.log"), index = 2)
formatter_data_frame <- function(df, ...) {
     pander::pander_return(df, style = 'simple')
}
log_formatter(formatter_data_frame)
log_formatter(formatter_glue)

## 
is_empty <- function(contrast) {
  return(length(contrast) == 0 || (length(contrast) == 1 && contrast == ''))
}

detect_file_in_folder <- function(folder_path, pattern) { 
          file <- list.files(folder_path, pattern = pattern); if(length(file) > 0) file[1] else "" }

if(is_empty(params$contrast) ){
  stop("Check parameter file: contrast is not set" )
}


if (all(params$comparison_label == '') ){
  stop("Check parameter file: comparisons label is not set" )
}

if (all(params$comparisons == '') ){
  stop("Check parameter file: comparisons is not set" )
}

if ( ! params$aggr_method %in% c('robustSummary','medianPolish','colMeans','colMedians')) {
  stop('normalization method not allowed.  Methods allowed are : [ robustSummary, medianPolish, colMeans ]' )
}
if (params$aggr_method  == 'medianPolish'){
  aggr_method_f <- MsCoreUtils::medianPolish
}
if (params$aggr_method  == 'RobustSummary'){
  aggr_method_f <- RobustSummary
}
if (params$aggr_method  == 'colMeans'){
  aggr_method_f <- base::colMeans
}
if (params$aggr_method  == 'colMeadians'){
  aggr_method_f <- matrixStats:::colMedians
}

if ( ! params$normalization %in% c("sum", "max", "center.mean", "center.median", "div.mean", "div.median", "diff.meda", "quantiles","quantiles.robust","vsn") ){ 
   stop('normalization method not allowed. Methods allowed are : [ sum, max, center.mean, center.median, div.mean, div.median, diff.meda, quantiles,quantiles.robust,vsn]' )
}

if (params$filtering_contaminant){
   if (params$contaminat_str == ""){
     stop("Please provide a string to filter the contaminats proteins (e.g.: Cont_,Contaminants,..)")
   } 
}

if (params$filtPerGroup == ''){
      label_filtering <- ' among all the samples.'
      }
if (params$filtPerGroup == 'all'){
      label_filtering <- ' in each group.'
      }
if (params$filtPerGroup == 'at_least_one'){
      label_filtering <- ' in at least one group.'
}


# make a subfolder for each contrast
if (!dir.exists(file.path(params$folder_prj))){
  dir.create(file.path( params$folder_prj),recursive = TRUE)
}
 dir.create(file.path( params$folder_prj, "Result"),recursive = TRUE)
 min_col_need_design <- c("Sample","Run", "Group", "Replicate")


 theme_set(theme_custom_vis())
 
 st=format(Sys.time(), "%Y%m%d_%H%M")

write(as.yaml(params), file =  file.path( params$folder_prj, paste0("DIAReport_parameter_",st  , ".yaml")))

```

## Import data

Reading DIA data from **`r basename(params$input_file)`** using *`r params$quantitative_features`* from DIA-NN as quantitative features.


`r if (! params$mbr) 'All the precursors with *Global Q-value and Global PG Q-value < 0.01* are used for the analysis' else 'All the precursors with *Lib Q-value and Lib PG Qvalue < 0.01* are used for the analysis'`

```{r import data}

  if (params$DIANN_ver2){
     log_info('Reading Data PARQUET format...')
     data <- read_parquet(params$input_file)
     lst_wide_columns <- c('Run', 'Precursor.Id', 'Modified.Sequence', 'Stripped.Sequence', 'Protein.Group', 'Protein.Ids', 'Protein.Names', 'Genes', 'Proteotypic')
     
  }else{
     log_info('Reading Data TSV format...')
     data <- read.csv(params$input_file,sep='\t')
     lst_wide_columns <- c('Run', 'Precursor.Id', 'Modified.Sequence', 'Stripped.Sequence', 'Protein.Group', 'Protein.Ids', 'Protein.Names', 'Genes', 'Proteotypic','First.Protein.Description')

  }
 
  ## deal csv  with different separated values 
  L <- readLines(params$design_file, n = 1)
  if (grepl(";", L)) design <- read.csv2(params$design_file) else design <- read.csv(params$design_file)

  checkDIANN <- check_DIANN_report(data, q_feature = params$quantitative_features)
    
  if (checkDIANN$status == 1 ) {
    stop(checkDIANN$error)
    } 
 
   
  dfMsqrob <- dfToWideMsqrob_20( data, precursorquan = params$quantitative_features, mbr = params$mbr, wide_colums = lst_wide_columns)

  
  if (params$DIANN_ver2){
    log_info('Retrieve Protein description (ver > 2.0)...')
    
    check_protein_desc <-  detect_file_in_folder(file.path(dirname(params$input_file)),
                                                  'report.protein_description.tsv')
    log_info(check_protein_desc)
    log_info( file.path(dirname(params$input_file),check_protein_desc))
    if (check_protein_desc != ""){
      prt_desc <- read.csv( file.path(dirname(params$input_file),check_protein_desc),sep="\t"  )
       
    }else{
       stop(paste0('file ','report.protein_description.tsv',  'not found in folder ->',
       	            dirname(params$input_file),'\n',' Check your input path !' ))
    }

    
    dfMsqrob<- dfMsqrob %>% mutate(app = sapply(str_split(Protein.Names, ";"), function(x) ifelse(length(x) > 0, x[1], NA))) %>% 
       left_join( prt_desc %>% select(Protein.Name,Description) , 
                  join_by(app == Protein.Name)) %>% 
      select(- app) %>% rename( First.Protein.Description =  Description   ) # filter(is.na(Description)) 
    }
  

  diann_colname <- c("Precursor.Id" , "Modified.Sequence","Stripped.Sequence","Protein.Group",
  "Protein.Ids","Protein.Names","Genes","Proteotypic","First.Protein.Description")
  
result_qfeat <-  import2_qfeature (dfMsqrob, design, params, min_col_need_design, diann_colname = diann_colname )

if  (result_qfeat$status == 1){
   stop(result_qfeat$error)
}else{
  
  pe <- result_qfeat$q_feat
}
  

```

## Filtering steps

```{r preparation }

  


  check_confounder_PCA<-checkConfounder(confounder= unique(unlist(str_split(params$PCA_comparison, "-"))), colsDesign=colnames(colData(pe)))
if (check_confounder_PCA$status == 1){
    stop(check_confounder_PCA$error)
}

pe <- add_rowdata_detection(pe,  design , assay = 'precursor' )


# rowData(pe[["precursor"]])$nNonZero <- pe[["precursor"]] %>%
#   assay %>%
#   is.na %>%
#   not %>%
#   rowSums
# 
# rowData(pe[["precursor"]])$pep_per_prot <-
#   left_join(rowData(pe[["precursor"]]) %>% as.data.frame %>% dplyr::select(Protein.Ids),
#             rowData(pe[["precursor"]]) %>% as.data.frame %>% dplyr::group_by(Protein.Ids) %>%
#               summarise(pep_per_prot = length(unique(Stripped.Sequence))))$pep_per_prot
# 
# ## add statistics per group 
# group_val <- design %>% distinct(Group) %>% 
#               arrange(Group) %>%  pull()
# group_size <- design  %>% group_by(Group) %>% summarise(n_=n()) %>% 
#               arrange(Group) %>%  pull(n_)
# names(group_size) <- group_val
# 
# 
# 
# for ( v in group_val){
#     log_info(v)
#     val = paste0('Zero',v)
#     rowData(pe[["precursor"]])[[val]] <- assay(pe[["precursor"]])[,rownames(colData(pe)[colData(pe)$Group == v,]),drop = FALSE]  %>% is.na  %>% rowSums()
#     ## 
#     val_due = paste0('perc',v)
#     rowData(pe[["precursor"]])[[val_due]] <- ((group_size[[v]] - rowData(pe[["precursor"]])[[val]] ) / group_size[[v]]) *  100
# }

orig_nFeat <- dim(pe[['precursor']])[1]
```

The filtering steps applied are:

-   Keeping only precursors with more than \>= **`r params$nNonZero`%** of *non missing* values, `r label_filtering`
-   `r if (params$Proteotypic) 'Selecting only proteotypic peptides' else 'Selecting both not and proteotypic peptides'`
-   Include proteins with at least **`r params$pep_per_prot` peptides**
-   `r if (params$filtering_contaminant) 'Contaminant proteins are filtered ' else 'Contaminant proteins are not filtered'`




```{r pre_processing}

size <- dim(colData(pe))[1]


pe_ <- filteringNA_qfeat_proteinwf <- function(pe , parameters, design){

# 
# group_val <- design %>% distinct(Group) %>% 
#       arrange(Group) %>%  pull()
#     group_size <- design  %>% group_by(Group) %>% summarise(n_=n()) %>% 
#       arrange(Group) %>%  pull(n_)
#     names(group_size) <- group_val
# 
# ## filtering per group
# 
#   if (params$filtPerGroup != '') {
#       log_info('filtering per group')
#     
#     ## or 
#    if (params$filtPerGroup == 'at_least_one') { 
#        log_info('filtering per group criteria: in at least one group ')
#        perc_group_val <- paste0("perc", group_val)
#        formula_condition <- paste(paste0(perc_group_val, " >= ", params$nNonZero), collapse = " | ")
#        dynamic_formula <- as.formula(paste0("~ ", formula_condition))
#    }
#     # and 
#     if (params$filtPerGroup == 'all') {
#         log_info('filtering per group criteria: for all the groups ')
#         perc_group_val <- paste0("perc", group_val)
#        formula_condition <- paste(paste0(perc_group_val, " >= ", params$nNonZero), collapse = " & ")
#        dynamic_formula <- as.formula(paste0("~ ", formula_condition))
#        }
#    
#    pe <-  filterFeatures(pe,dynamic_formula)
#     
# 
#    }else{
#         log_info('filtering across all samples')
#        pe <- filterFeatures(pe, ~ nNonZero >= round(size  * ( params$nNonZero / 100))   )
# 
#   }  

# Proteotypic features
# if (params$Proteotypic){
#   log_info('Proteotypic filtering')
# 
#   pe <- filterFeatures(pe, ~ Proteotypic == 1)
# 
# }
# 
# ## filtering out contaminats
# 
# if (params$filtering_contaminant){
#     log_info('Filtering contaminants')
# 
#   pe <- filterFeatures ( pe ,VariableFilter("Protein.Ids", params$contaminat_str, "contains", not=TRUE))
# 
# }
# 
# # At least 2 peptides per protein across all groups
# log_info('Filtering only n peptides per proteins')
# 
# pe <- filterFeatures(pe, ~ pep_per_prot > params$pep_per_prot)


log_info('Intensity log tranformation')

pe <- logTransform(pe, base = 2, i = "precursor",
                   name = "precursorLog")
log_info('Normalization')


pe <- QFeatures::normalize(pe,  method = params$normalization, i = "precursorLog",
                name = "precursorNorm")
#MsCoreUtils::medianPolish()
log_info('Summarization at protein level')

pe <- aggregateFeatures(pe, i = "precursorNorm",
                        fcol = "Protein.Ids",
                        name = "proteinRS",
                        fun = aggr_method_f,
                        na.rm = TRUE)

log_info('Computing protein level % not missing')

pe <- add_rowdata_detection(pe,  design , assay = 'proteinRS' )



```

The initial number of precursor features  was **`r orig_nFeat`**, after filtering  **`r dim(pe[['precursor']])[1]`** features are retained.

```{r adding_ensembl_annotation}
if ( ! params$ensembl_annotation == ''){
  ## read it 
  ensembl_db_table <- read.csv(params$ensembl_annotation)
  ## make it simple 
  
  ensembl_db_table <- ensembl_db_table %>%  distinct(ensembl_gene_id , .keep_all=TRUE)  %>% 
      dplyr::select( params$ensembl_col    ) %>% distinct( hgnc_symbol, .keep_all=TRUE)
  app <- as.data.frame(rowData(pe[['proteinRS']]))  %>%  left_join( ensembl_db_table , join_by(Genes ==  hgnc_symbol)) 
  
  for (i in  1:(length( params$ensembl_col) -1) ){
     ann_col <- params$ensembl_col[i]
     
     rowData(pe[['proteinRS']])[[ann_col]] <- app[[ann_col]]
  }
 
}

```

## Confounder Analysis
```{r confounder_value_analysis}
log_info('Confounder Analysis')
cData<-as.data.frame(colData(pe))
list_numeric <- list()
list_character <- list()

if ( ! is_empty(params$confounder_list) ) {
  for  (col in params$confounder_list){
     if  (is.factor (cData[[col]])) {
       #log_info(paste0('Character - ', col))
       list_character <- append(list_character, col)
        
     }else {
       #log_info(paste0('Numeric - ', col))
       list_numeric <- append(list_numeric, col)
     }  
  }

  if (length(list_numeric) != 0) {
  
    cData_n<-cData%>%pivot_longer(cols=unlist(list_numeric), names_to="Numeric_Variables", values_to="Value")

		           boxplot_num <-ggplot(cData_n, aes(x=Numeric_Variables, y=Value, fill=Group))+geom_boxplot()
  print(boxplot_num)
               pdf(file= file.path(params$folder_prj,"Result",paste0('Boxplot_num_confounder.pdf')),paper="a4")
               boxplot_num <-ggplot(cData_n, aes(x=Numeric_Variables, y=Value, fill=Group))+geom_boxplot()
  print(boxplot_num)
      invisible(dev.off())
  } 
  
  for (col_ch in unlist(list_character)){
    #log_info(col_ch)
    boxplot_ch<- ggplot(cData, aes(x={{col_ch}}, fill=Group))+geom_bar()+ggtitle( {{col_ch}} )
    print(boxplot_ch)
          pdf(file= file.path(params$folder_prj,"Result",paste0('Boxplot_ch_confounder.pdf')),paper="a4")
         boxplot_ch<- ggplot(cData, aes(x={{col_ch}}, fill=Group))+geom_bar()+ggtitle( {{col_ch}} )
    print(boxplot_ch)  
     
      invisible(dev.off())
  }

}


```


## Normalization

Normalization of the precursor intensities are:

-   Log transformed and then normalized using **`r params$normalization`** approach.
-   Precursors/peptides are summarizated at protein levele using the **`r params$aggr_method`** approach.

After all the step, the quantitaive features used in the downstream analysis are:

- Number of precursor identified/quantified is : **`{r} dim(assay(pe[['precursor']]))[1]`**
- Number of protein identified/quantified is : **`{r} dim(assay(pe[['proteinRS']]))[1]`**


```{r normalization_plot}
par(mfrow=c(1,3))
limma::plotDensities(assay(pe[["precursorLog"]]),legend=FALSE,main='Log-Transform')
limma::plotDensities(assay(pe[["precursorNorm"]]),legend=FALSE,main='Normalized') 
limma::plotDensities(assay(pe[["proteinRS"]]),legend=FALSE,main='Protein Summarization' )

```


## Data Missing Analysis

The plot shows the completeness of the experiments at precursor and summarized level. On the the x-axis peptides/proteins are ordered based on the increasing NA content.

```{r missing_val_analysis}
peptidemissingness <- MSnbase::plotNA(assay(pe[["precursorNorm"]])) +
  xlab("Precursor index (ordered by data completeness)") +
  theme_bw() +
  theme(legend.position = "none") +
  theme(strip.text.x = element_blank(),
        strip.background = element_rect(colour="white", fill="white"),
        legend.position=c(.35,.15)
  ) +   ggtitle("Missingness - Precursor level" )

#peptidemissingness

proteinmissingness <- MSnbase::plotNA(assay(pe[["proteinRS"]])) +
  xlab("Protein index (ordered by data completeness)") +
  theme_bw() +
  theme(legend.position = "none") +
  theme(strip.text.x = element_blank(),
        strip.background = element_rect(colour="white", fill="white"),
        legend.position=c(.35,.15)
  ) +   ggtitle("Missingness - Protein Level")
#proteinmissingness
#missingness <- gridExtra::grid.arrange(peptidemissingness, proteinmissingness, nrow = 1)
#print(missingness)

```


## Completeness Analysis 

This section analyzes the completeness of protein readouts across all samples within each group. For each quantified protein, a completeness percentage is calculated based on the proportion of missing values across all samples in the group. These percentages are compared between the groups in the lower diagonal. On the diagonal, a histogram illustrates the distribution of detection percentages for each group

```{r completeness_analysis}
log_info('Completness Analysis...')
   custom_lower <- function(data, mapping, ...) {
    
      x_data <- data[[as_label(mapping$x)]]
      y_data <- data[[as_label(mapping$x)]]
     
      
      # Calculate mean and standard deviation
      mean_x <- mean(x_data, na.rm = TRUE)
      mean_y <- mean(y_data, na.rm = TRUE)
      sd_x <- sd(x_data, na.rm = TRUE)
      sd_y <- sd(y_data, na.rm = TRUE)
      
      ggplot(data = data, mapping = mapping) +
      geom_jitter(  size=1,alpha= 0.8)+
      geom_abline(slope = 1, intercept = mean_y - mean_x, color = 'blue')+
      geom_abline(slope = 1, intercept = mean_y - mean_x + sd_y + sd_x, color = 'red', linetype = "dashed") +       # Mean + sigma line
        geom_abline(slope = 1, intercept = mean_y - mean_x - sd_y - sd_x, color = 'red', linetype = "dashed")       # Mean - sigma line
      
      
    }
    
    
    custom_diag <- function(data, mapping, ...) {
      ggplot(data = data, mapping = mapping) + 
        geom_histogram(binwidth = 10, fill = "blue", color = "black") 
    }
    
    
    custom_diag <- function(data, mapping, ...) {
      ggplot(data = data, mapping = mapping) + 
        geom_histogram(binwidth = 10, fill = "blue", color = "black") 
    }
    
    
    as.data.frame(rowData(pe[['proteinRS']])) %>% 
      select(matches("^perc"))  %>% 
      ggpairs(
               upper = list(continuous = wrap("blank")),
              diag = list(continuous = custom_diag),
      lower = list(continuous = custom_lower),
      xlab = "percentage of detection",
      ylab = "percentage of detection"
      )
  

```


## PCA

```{r pca_plot}
#| results: asis

# to do make in for loop for a set of variable
log_info('PCA Analysis...')
var_topca <- params$PCA_comparison

l_pca <- generate_pca_plots(var_topca, pe, params, layer ='proteinRS' )
if (length(l_pca) == 0){
  stop('Error during PCA generation ')
}
#ggplotly(  l_pca[[1]])

#htmltools::tagList(l_pca)
# saveRDS(l_pca, 'pcaPlot.Rds')
 for (pca_plot in l_pca ){
   
  print(htmltools::tagList(ggplotly(pca_plot)))
   #ggplotly(pca_plot)
 }

```

## DE Analysis

Using MSqRob2 with the following formula `r params$formula` with ridge regression disabled

```{r build_model}
log_info('DE Analysis with Msqrob..')
pe <- msqrob(object = pe, i = "proteinRS", 
             formula = as.formula( params$formula)  ,ridge = FALSE, overwrite = TRUE)


contrast_list <- paste0(params$comparisons, "=0")


coef <- rowData(pe[["proteinRS"]])$msqrobModels[[1]] %>% getCoef %>% names

    if (is.null(coef)) {
          coef <- rowData(pe[["proteinRS"]])$msqrobModels[[2]] %>% getCoef %>% names
            #log_info(getCoef(rowData(pe[['proteinRS']])$msqrobModels[[2]]))
          if (is.null(coef)){
            stop ('Model is not able to converge')
          }
          getCoef(rowData(pe[['proteinRS']])$msqrobModels[[2]])


    }else{
      #log_info(getCoef(rowData(pe[['proteinRS']])$msqrobModels[[1]]))
      ## print coefficent 
      getCoef(rowData(pe[['proteinRS']])$msqrobModels[[1]])

    }

## to be removed 
#coef <-  unique(unlist(strsplit(params$comparisons,' - ',fixed=T)))
## coef from the model itself
#coef <- rowData(pe[["proteinRS"]])$msqrobModels[[1]] %>% getCoef %>% names


L <- makeContrast(contrast_list, parameterNames = coef)

# L <- makeContrast(c( "ridgeGroupB - ridgeGroupA = 0"), parameterNames = c("ridgeGroupB","ridgeGroupA"))
#rownames(L) <- paste0("ridge",rownames(L))


pe <- hypothesisTest(object = pe, i = "proteinRS", contrast = L , overwrite=TRUE)


test <-  lapply(params$comparisons, dep_volcano, data= pe,  p=params )

# params$comparison
names(test) <-  params$comparison_label

log_info('Saving pe object (qfeatures) as Rds file.')
saveRDS(pe, file = file.path( params$folder_prj, "Result","qfeat_experiment.Rds"))


if (!dir.exists(file.path(params$folder_prj, "Result","Contrast"))){
  dir.create(file.path( params$folder_prj, "Result","Contrast"), recursive = TRUE)
}
# old one params$comparisons
for (a in   params$comparison_label ){
  log_info(a)
  log_info(check_and_substitute_forbidden_chars(a))
  dir.create(file.path(params$folder_prj,"Result","Contrast",check_and_substitute_forbidden_chars(a)), recursive = TRUE,showWarnings = TRUE)
}



```

## QC plots

::: panel-tabset
```{r QC_plot}
#| results: asis
  for (cmp in names(test)) {
    cmp_label <- gsub(params$contrast,'',cmp )
    cat( '##', paste(cmp_label,' P-values distribution 
                     ',sep=' '), '\n\n')
    
    render_child(data = test[[cmp]], 
                 pe =pe , 
                 label= cmp ,
                 path=file.path(params$folder_prj, "Result","Contrast",check_and_substitute_forbidden_chars(cmp)), template='_templatePval_distribution.Rmd') 
    cat('\n\n')
    
  }

```
:::

## Comparison

::: panel-tabset
```{r comparison}
#| results: asis

log_info('Running Analysis for each comparison..')

  for (cmp in names(test)) {
        log_info( paste0('  --  ',cmp)) 

    cmp_label <- gsub(params$contrast,'',cmp )
    cat( '## ', paste(cmp_label,sep=' '), '\n\n')
    cat ('This volcano plot summarizes the differential expression landscape in the comparison between the two groups \n')
    
    
    render_child(data = test[[cmp]],
                 pe =pe,
                 label = cmp , 
                 path=file.path(params$folder_prj, "Result","Contrast",check_and_substitute_forbidden_chars(cmp)),  
                 template='_templateContrast.Rmd') 
    cat('\n\n')
    
   

    if ( dim(test[[cmp]]$toptable %>% filter(differential_expressed %in% c('UP','DOWN')) ) [1] <= 5   ){
      cat(' Not enough DE proteins for heatmap visualization \n')
    }else{
      
    #filt_val <- unlist(str_split(substring(gsub(params$contrast, ' ', (gsub(' - ','',cmp))), 2),' ' ))
    #print(filt_val)
    
    contrast_values_paresed <- parse_comparison(cmp, params$contrast,pe)

    filt_val <- select_samples_comparison(contrast_values_paresed, pe , params$contrast)
    
    render_child(data = test[[cmp]],
                   pe = pe,
                   sample_rel = filt_val,
                 label = cmp,
                 path= file.path(  params$folder_prj, "Result","Contrast",check_and_substitute_forbidden_chars(cmp)),
                 template='_templateHeatMap.Rmd')
    }
    cat('\n\n')
    cat('\n\n')
    cat('\n\n')
    cat('These bar plots summarizes the number of significantly upregulated/downregulated number of proteins based on different adjusted p-values (selected adjusted p-values are 0.001, 0.01, 0.05, and 0.1 - see facet headers) and log2 fold-change thresholds (on the x-axis) used to define the significance levels.')
   
    render_child(data = test[[cmp]],
                 pe =pe,
                 label = cmp,
                 path= file.path( params$folder_prj, "Result","Contrast",check_and_substitute_forbidden_chars(cmp)),
                 template='_templateBarPlot.Rmd')
    cat('\n\n')
    
    
  }

```
:::



## Summary of the DE proteins

In the table you can find a summaries of the number of DE proteins found in all the comparisons. The overlapping of DE proteins among the comparisons is also showed using an upset plot.

```{r upset plot}
log_info('DE Summaries computation...')
listInput <- list()
df_summary = as.data.frame(matrix(ncol=3, nrow=0))
i <- 1 
for (cmp in names(test)){
  cmp_label <- gsub(params$contrast,'',cmp )
  total =  dim(test[[cmp]]$toptable %>% filter(differential_expressed =='UP' |  differential_expressed =='DOWN') )[1]
  up  = dim(test[[cmp]]$toptable %>% filter(differential_expressed =='UP') )[1]
  down  =   dim(test[[cmp]]$toptable %>% filter(differential_expressed =='DOWN') )[1]
  log_info(cmp)
  log_info(total)
  #df_summary[[cmp_label]]  <- c(total, up,down)
  df_summary[i,1]  <- total
  df_summary[i,2]  <- up
  df_summary[i,3]  <- down

  if (total >  0){
    listInput[[cmp_label]] <- test[[cmp]]$toptable %>% filter(differential_expressed =='UP' |  differential_expressed =='DOWN') %>% pull(Protein.Names)
  }
  i <- i+1
}
rownames(df_summary) <- names(test) 
colnames(df_summary) <- c('Total DE','Up DE','Down DE')
    ## visual table summary
    DT::datatable( df_summary  ,
             extensions = c('FixedColumns', 'Scroller'),
             options = list(fixedColumns = TRUE, scrollY = 400, scrollX = TRUE,
                            scroller = TRUE, dom = 'Bfrtip', autoWidth = TRUE
             ),
     ) %>% formatStyle(
   columns= c('Total DE','Up DE','Down DE'),
   textAlign = 'center')
    # if there's 1 or 2 elements inlistInput and one of them is empty, skip uspset plot
    #log_info({length(listInput)})
 if (length(listInput)>=2 ) {
    #heuristic_interaction <- round(mean(sapply(p, length)) /2 )
    print (upset(fromList(listInput), order.by = "freq", nsets = length(listInput) ) )
     
      pdf(file= paste0(file.path( params$folder_prj,"Result","Contrast"),'Upset_plot.pdf'),paper="a4")
       
     print( upset(fromList(listInput),  nsets = length(listInput) , order.by = "freq",text.scale= 1,
                  empty.intersections = "on")) 
     
      invisible(dev.off())
      
      
 
}
    

```

## Session Info

```{r session_info }
sessionInfo()
```
